Union and intersection contracts are hard, actually

### 2021-06-09 paper.tex
Fix citation
Pass on intersection contracts

Intersections ($\cap$) are less prevalent. An intersection is a value that belongs to
two types at the same time. An example is a number literal \nickel{1}
which could be accepted as both an \nickel{Int} and a \nickel{Float} without
having to perform conversions.
In that case, a value of type \nickel{Int @& Float} can be used as a value of
both type \nickel{Int} and \nickel{Float}.

% \label{fig:intersection-subytping-ex}
% \begin{lstlisting}[title={Value subtyping}]
% 1 : Int @& Float
% \end{lstlisting}

The real usability of intersections shows up when intersecting
functions, since it can be used to encode overloading.
For instance, take a look at Figure
\ref{fig:intersection-overloading-ex}, where the function
\nickel{duplicate} works both as a function to duplicate
arrays, as well as a function to duplicate strings.
This is particular useful when using unions, since it's a
good way to express that a function can deal with different
shapes of data.
For instance, on the same figure, you can use \nickel{duplicate}
over some value with type \nickel{[String] @| String}, and know
that it will work correctly.

\begin{figure}[h]
\begin{lstlisting}[language=nickel]
let duplicate | ([String] -> [String]) @& (String -> String) =
  fun x => x ++ x in
let text | [String] @| String = ... in
duplicate text
\end{lstlisting}
\caption{Duplicating an array of Strings or a String}
\label{fig:intersection-overloading-ex}
\end{figure}

% \label{fig:intersection-overloading-ex}
% \begin{lstlisting}[title={Overloading}]
% incr: (Int -> Int) @& (Float -> Float) =
%   fun x => x + 1
% \end{lstlisting}

They can also be used to combine dictionaries in an object-oriented fashion. For instance, in Figure \ref{fig:intersection-record}
two types are defined \nickel{Animal} and \nickel{Pet}, and
a variable that is compatible with both types is declared, with
type \nickel{Animal @& Pet}.

@@

% They can also be used to combine constraints (such as traits or interfaces),as
% done in Scala 3 (Figure~\ref{fig:intersection-mixins-ex}).

% \label{fig:intersection-mixins-ex}
% \begin{lstlisting}[language=Scala,title={Structural mixins}]
% Comparator: {cmp: T -> T -> Bool}
% Shower: {show: T -> String}

% showSmaller : (dict: Comparator & Shower) -> T -> T -> String
% \end{lstlisting}

Finally, intersection naturally show up as the dual of unions.
Thus, in a system
featuring unions, a union in a negative positions is a form of intersection
(Figure~\ref{fig:intersection-dual}).

\begin{figure}[h]
\begin{lstlisting}[language = nickel]
f : (Num @| Str) -> Bool
(*@ $\simeq$ @*) f : (Num -> Bool) @& (Str -> Bool)
\end{lstlisting}
\caption{Intersection as the dual of unions}
\label{fig:intersection-dual}
\end{figure}

\resolved{Either here or the next section, show examples of unintuitive behaviour
of U and I. For example, I is not jos composition of contracts.
And union has to be treated as any one of the options.
Kindo of resolved for unions}
% Things to settle: how to motivate them, how to compare them to ADTs. Purely
% dynamic approach, or typed approach too?

@@

Enforcing pre- and post-conditions at runtime is a widely established practice.
In their foundational paper\cite{FindlerFelleisenHOContracts}, Findler and
Felleisen introduce higher-order contracts, a principled approach to run-time
assertion checking that nicely supports functions. They introduce the notion of
blame, which is crucial to good error reporting. It became apparent later that
their contracts are closely related to the type casts introduced by gradual
typing, excluding blame: both \cite{FindlerMultiLang} and
\cite{FelleisenInterLang} see the value of contracts as a safe interface between
typed and untyped code. In \cite{WellTypedBlamed}, the authors precisely
introduce a system integrating gradual typing with contracts \textit{à la
Findler \& Felleisen}. Nickel adopts a similar type system, with both statically
typed terms, dynamically typed terms, and first-class contracts. Higher-order
contracts are the basis of the work~\cite{KeilThiemannUnionIntersection}
and~\ref{RootCauseOfBlame} that this paper studied in detail.

>>>>

An intersection type \nickel{A @& B} validates values which are both
valid for contract \nickel{A} and contract \nickel{B}.

Intersection contracts (and types) are probably less prevalent than
union in practical type systems. However, a function from a union is
equivalent to an intersection. That is \nickel{(A @| B) ->
  C}$\simeq$\nickel{(A -> C) @& (B -> C)}. So in a system with
functions and unions, intersections are already morally present (and,
for that matter, in a system with functions and intersections, unions
are morally present). Some
of our examples in Sections~\ref{sec:issues-sem}
and~\ref{sec:issues-literature} are better expressed in terms of
intersections, so it's best to include them.

Yet intersections are useful in their own right. They can be used to
combine dictionaries in the style of object-oriented inheritance. For instance, in
Figure \ref{fig:intersection-record} two types are defined
\nickel{Animal} and \nickel{Pet}, and a variable that is compatible
with both types is declared, with type \nickel{Animal @& Pet}. This
particular application is supported, for instance, by TypeScript.

@@

Another application of intersections shows up when intersecting
functions: it can be used to encode overloading.
For instance, take a look at Figure~\ref{fig:intersection-overloading-ex}, where the function
\nickel{duplicate} works both as a function to duplicate
arrays, as well as a function to duplicate strings.
This is particularly useful when using unions, since it's a
good way to express that a function can deal with different
shapes of data.
For instance, in the same figure, \nickel{duplicate} is (correctly) called
on a value of type \nickel{[String] @| String}.

\begin{figure}[h]
\begin{lstlisting}[language=nickel]
let duplicate
    | (List String -> List String)
    @& (String -> String) =
  fun x => x ++ x in
let text | [String] @| String = ... in
duplicate text
\end{lstlisting}
\caption{Duplicating an array of Strings or a String}
\label{fig:intersection-overloading-ex}
\end{figure}

@@

Enforcing pre- and post-conditions at runtime is a widely established practice.
In their foundational paper\cite{FindlerFelleisenHOContracts}, Findler and
Felleisen introduce higher-order contracts, a principled approach to run-time
assertion checking that nicely supports functions. They introduce the notion of
blame, which is crucial to good error reporting. It became apparent later that
their contracts are closely related to the type casts introduced by gradual
typing, excluding blame: both \cite{FindlerMultiLang} and
\cite{FelleisenInterLang} see the value of contracts as a safe interface between
typed and untyped code. In \cite{WellTypedBlamed}, the authors precisely
introduce a system integrating gradual typing with contracts \textit{à la
Findler \& Felleisen}. Nickel adopts a similar type system, with both statically
typed terms, dynamically typed terms, and first-class contracts. Higher-order
contracts are the basis of the work~\cite{KeilThiemannUnionIntersection}
and~\cite{RootCauseOfBlame} that this paper explorted extensively.
